itle: "Reproducible Research"
output: html_document
---
### Loading and processing the data
use the necessary library and read the data
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE) 
library(readr)
library(knitr)

activity <- read_csv("C:/Users/IAS User/Desktop/activity.csv")
# see the summary of the data
summary(activity)
```

```{r}
# Remove the NA values 
good_act <- activity[complete.cases(activity), ]

```
### Histogram of the total number of steps taken each day
For this part, first I will calculate the total number of steps taken each day and then create a histogram
```{r}
# calculate the total number of steps taken each day
steps_per_day <- aggregate(steps ~ date, good_act,sum)
# create the histogram 
hist(steps_per_day$steps, main = "Histogram of the total number of steps taken each day", col = "green", xlab = "Steps per day")
```

### calculate the mean and median number of steps taken each day
```{r}
# calculate the mean
mean(steps_per_day$steps)
```

```{r}
# calculate the median
median(steps_per_day$steps)
```
### Time series plot of the avarage number of the steps taken
1. Make a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}

# calculate average steps per interval for all days

avg_steps_per_interval <- aggregate(steps ~ interval, good_act, mean)

# plot the time series
plot(avg_steps_per_interval$interval, avg_steps_per_interval$steps, type='l', col=3, main="The average number of steps taken by interval", xlab=" Time intervals", ylab="Average number of steps")
```

### The 5 mintue interval that, on average, contains the max number of days 
```{r}
max_interval<- which.max(avg_steps_per_interval$steps)
avg_steps_per_interval[max_interval,]$interval
```
### Code to describe and show a strategy for imputing missing data
1.Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

3.Create a new dataset that is equal to the original dataset but with the missing data filled in.


```{r}
# Calculate the missing values
bad_act  <-activity[!complete.cases(activity),]
nrow(bad_act)
```


```{r}
# Loop through all the rows and find which step has NA in that row.Then identify the avg steps for that interval in avg_steps_per_interval and substitute the NA value with that value 

for (i in 1:nrow(activity)) {
    if(is.na(activity$steps[i])) {
        val <- avg_steps_per_interval$steps[which(avg_steps_per_interval$interval == activity$interval[i])]
        activity$steps[i] <- val 
    }
}
```
### Histogram of the total number of steps taken each day after values are imputed

```{r}
# The steps per day with the imputed values
steps_per_day_impute <- aggregate(steps ~ date, activity, sum)
# Create a histogram 
hist(steps_per_day_impute$steps, main = "Histogram of total number of steps per day (IMPUTED)", xlab = "Steps per day", col="green")
```

```{r}
# Calculate the mean 
mean(steps_per_day_impute$steps)
```
```{r}
# calculate the median
median(steps_per_day_impute$steps)
```
### Panel plot comparing the average number of steps taken per 5 mintue interval across weekdays and weekends

```{r}
 week <- (activity$date)
 week <-weekdays(as.Date(week))
 class(week)
 week <- as.factor(week)
 table(week)
 act <- as.Date(week, "%Y,%m,%d")
 
 

```
